---
title: "01.2_PCA_analysis"
author: "Taiba"
date: "2023-04-23"
output: pdf_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor, tidymodels, factoextra, ggfortify, gridExtra, knitr)

```

```{r}
chemi_class<- read_csv("final_chemv5.csv") |>
  filter(compound != "PICLORAM" & compound != "METSULFURON" & compound != "TRICLOPYR")
```

```{r}
# pivot from long to wide format contains 3069 obs with 28 variables
chemi_wide <- chemi_class[c(1:4)] |>
  pivot_wider(names_from = compound, values_from = median_pest) |>
  select(-c(state, county_fips_code))|>
 mutate_at(vars(1:23), ~log10(.+0.0000001)) 

#replacing NA values due to missing information from original data from EPA
# replace all missing values with log(0.0000001) = -16.1181
chemi_wide_new <- chemi_wide |>
  mutate(across(everything(), ~replace(., is.na(.), -16.1181)))

# Calculate principal components
#scale: a logical value indicating whether the variables should be scaled to have unit variance before the analysis takes place
pca.ln <- prcomp(chemi_wide_new, scale = TRUE)
save(pca.ln, file = "pca_ln.rda")
```
########################################################################
########################################################################
        WORK FROM HERE

########################################################################
########################################################################

```{r}
#load model object
load("pca_ln.rda")
```

```{r}
eigenvalues_ln <- matrix(pca.ln$sdev^2) #eigenvalues
perc_variance <- round(100*matrix(pca.ln$sdev^2/sum(pca.ln$sdev^2)),1) #variance

#Summary table 
eigenvalues_ln <- cbind(1:24, eigenvalues_ln, perc_variance) 
colum_ln <- c("Principal Component", "Eigenvalues", "Percent Variance")
eigenvalues_ln <- kable(eigenvalues_ln, col.names = colum_ln)
eigenvalues_ln 
```

## Proportion of Variance Plots
The first two components explained 55% of the variance 
```{r PVE plots}
#Plots the proportion of variance explained by each component (scree plot)
pve.ln <- pca.ln$sdev^2/sum(pca.ln$sdev^2) #proportion of variance explain by each component
 
#log-transformed data
fviz_eig(pca.ln, main = "",
         xlab = "Principal component",
         ylim = c(0,50))

ggsave("pca_screeplot.tiff",
       width=6, height= 6, dpi=300)
```


## Data Visualization of eigenvectors w/ Log-Transformed PCA Results
Loadings are the weights that each chemical contribute to the component. Scores are the sum of loadings multiply by concentration of each chemical for each person. So you get a loading for each chemical in each component and also a total loading for each principal component (which is the sum of the chemical's loadings). You also get a score for each person (each observation) which is the sum of the scores of each chemical (loading*chemical concentration). So for each person (observation) you have a score for each principal component. Each principal component also has a score which is the sum of the scores within the principal component. 

```{r}
pca.ln.ld <- as.data.frame.matrix(pca.ln$rotation) ## rotation is the loadings variable within the pca output.
pca.ln.ld$chem <- row.names(pca.ln.ld)

#run compounds_classes from file 01_chemicals_sorting.Rmd
# Convert the list to a tibble and unnest the list
loadings_pca <- pca.ln.ld |>
  mutate(Group = case_when(
    chem %in% unlist(compound_classes) & chem %in% compound_classes$Herbicide ~ "Herbicide",
    chem %in% unlist(compound_classes) & chem %in% compound_classes$Insecticide ~ "Insecticide",
    chem %in% unlist(compound_classes) & chem %in% compound_classes$Fungicide ~ "Fungicide",
    TRUE ~ NA_character_
  ))

plot_loadings_pca <- loadings_pca |> 
  gather(key = "PC", value = "Loading", -chem, -Group) |> as.tibble()

################################################################
chem_order <- c("2,4-D", "ATRAZINE", "ACETOCHLOR","CLETHODIM", "DICAMBA","GLYPHOSATE", 
                                      "IMAZETHAPYR", "PARAQUAT", "PENDIMETHALIN","CHLORIMURON","METOLACHLOR",
                                      "METRIBUZIN", "NICOSULFURON","SETHOXYDIM", 
                                      "SIMAZINE","THIFENSULFURON","TRIBENURON METHYL", "TRICLOPYR", "TRIFLURALIN","CARBARYL",         "CHLORPYRIFOS", "CYHALOTHRIN-LAMBDA","DIMETHOATE", "ESFENVALERATE","PERMETHRIN","CHLOROTHALONIL","MANCOZEB","PROPICONAZOLE")

plot_loadings_pca |> 
  filter(PC %in% c("PC1", "PC2", "PC3")) |> 
  mutate(PC = as.factor(PC),
         PC = fct_recode(PC, "PC 1" = "PC1",
         "PC 2" = "PC2",
         "PC 3" = "PC3"),
         chem = factor(chem, levels = chem_order)) |> 
  ggplot(aes(x = chem, 
             y = Loading, 
             fill=Group)) + 
  geom_col() +
  facet_wrap(~ PC) + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90, hjust = 1),
        strip.background = element_rect(fill = "white")) +
  geom_hline(yintercept = 0, size = 0.2) +
  labs(x = "Chemicals",
       y = "Loadings")

ggsave("pca_loadings.tiff",
       width=10, height= 7, dpi=300)
```

## Principal Component Biplot w/ Log-Transformed PCA Results

```{r PCA biplot}
#Creates a biplot showing the loadings for principal component 1 and 2. 
autoplot(pca.ln, data = chemi_wide_new, size = 0.8, colour = 'blue', alpha = 0.5,
         loadings = TRUE, loadings.colour = 'orange',
         loadings.label = TRUE, loadings.label.repel = T, 
         loadings.label.size = 2.5, loadings.label.colour = 'black',
         main = "Principal Component Analysis Biplot")

ggsave("pca_biplot.tiff",
       width=8, height= 8, dpi=300)
```

## PCA index score using median magnitude of first 5 components(or those with eigen values>=1) using Naveen's approach. We extract the first 5 components using [, 1:5]. Then, for each observation, we calculated the magnitude by taking the square root of the sum of squares of the 5 component values. Finally, we estimate the median of these magnitudes and store it in the variable score.

```{r}
# Naveen - Median magnitude
#Extract the first 5 components
first_5_components <- pca.ln$x[, 1:5]

# Calculate the magnitude of each observation's first 5 components
component_magnitudes <- apply(first_5_components, 1, function(x) sqrt(sum(x^2))) |>
  as.data.frame() |>
  rename_at(1, ~"med_magni")
```

```{r}
### Method 2- PCA analysis Dr.Beseler approach
#weighted scores
factor_loadings <- pca.ln$rotation

# Extract the scores with rotation
scores <- pca.ln$x

# multiplication of pesticide value with loading values by each county
result_test <- tibble(result = scores * chemi_wide_new)

#extract the multiplied values -> perform row sum of multiplied values -> retain only the index variable
result_test_df<- pull(result_test, result) |>
  mutate(chem_index= rowSums(across(everything())))|>
  select(chem_index)

#combine chem_index and median_magnitude into a dataframe
new_index_vars<- cbind(result_test_df, component_magnitudes)

new_index_vars_withpc1<- cbind(new_index_vars, pca.ln$x[,1])
```

```{r}
## Save the summed score as a CSV file
write_csv(new_index_vars_withpc1, "index_scores.csv")

```

#combine index variables with chemicals applied data
```{r}
chem_mapping <- chemi_class[c(1:4)] |>
  pivot_wider(names_from = compound, values_from = median_pest) |>
  mutate_at(vars(3:29), ~log10(.+0.0000001)) |>
  mutate(across(everything(), ~replace(., is.na(.), -16.1181)))

# final file for mapping
dat_map<- cbind(chem_mapping, new_index_vars_withpc1)
write_csv(dat_map, "mapping_dat_final.csv")
```

