---
title: "01.4_SVI_PCA_Spatial_analysis"
author: "Taiba"
date: "2024-03-07"
output: pdf_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor, tidycensus, leafsync,ggspatial, mapview, rmarkdown, knitr, stringr) #ggspatial will get spatial info
# census_api_key("60cbacc5c89ceb76d50a175907d1a9af7a7f3a1b", install = TRUE)
```

#Code to check ACS Survey variables
```{r}
#load variables
vars <- load_variables(2019, "acs5") #Loads all the variables

vars <- load_variables(2019, "acs5/profile")

#variables used for Social vulnerability
#Minority race variables
  #B02001_003 - Estimate!!Total:!!Black or African American alone
  #B02001_004 - Estimate!!Total:!!American Indian and Alaska Native alone
  #B02001_006 - Estimate!!Total:!!Native Hawaiian and Other Pacific Islanders
  #B03002_012 - Estimate!!Total:!!Hispanic or Latino

# No diploma/less than high school
#B15003_016 - Estimate!!Total:!!12th grade, no diploma, EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER
#B15003_013- Estimate!!Total:!!9th grade EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER
#B15003_014- Estimate!!Total:!!10th grade EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER
#B15003_015- Estimate!!Total:!!11th grade EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER

# Below 200% Federal poverty level
#C17002_002 Estimate!!Total:!!Under .50 RATIO OF INCOME TO POVERTY LEVEL IN THE PAST 12 months
#C17002_003 Estimate!!Total:!!.50 to .99 RATIO OF INCOME TO POVERTY LEVEL IN THE PAST 12 MONTHS

```

#SVI Variables for 48 US States
```{r}
#get list of variables from 2015-2019 census
acs_variable_list<- load_variables(2019, "acs5", cache = T)
svi_us<- get_acs(geography = "county",
                  state = c("al", "ar", "az", "ca", "co", "ct", "dc","de", "fl", "ga", "ia", "id",
                  "il", "in", "ks", "ky", "la", "ma", "md", "me", "mi", "mn", "mo", "ms",
                  "mt", "nc", "nd", "ne", "nh", "nj", "nm", "nv", "ny", "oh", "ok", "or",
                  "pa", "sc", "sd", "tn", "tx", "ut", "va", "vt", "wa", "wi", "wv", "wy", "ri"),
                  year=2019,
                  survey = "acs5",
                 variables = c(Black_tot= "B02001_003", ameri_ind_tot= "B02001_004", native_hawai_tot= "B02001_006", hisp_latino_tot= "B03002_012", no_diploma= "B15003_016", bel_10th= "B15003_013",bel_11th= "B15003_014", bel_12th= "B15003_015", bel_pov1= "C17002_002", bel_pov2= "C17002_003"),
                  geometry = TRUE,
                  output = "wide") %>% clean_names()

```


```{r}
# Separate county and state names
svi_us <- svi_us %>%
  separate(name, into = c("county", "state"), sep = ", ", remove = FALSE)

# Create new variable "minority" by adding the specified variables
#svi_us <- svi_us %>% 
 # mutate(minority = black_tot_e + ameri_ind_tot_e + native_hawai_tot_e + hisp_latino_tot_e) %>%
  
# Convert 'geoid' column to numeric
svi_us$geoid <- as.numeric(svi_us$geoid)

# Checking for missing values
missing <-svi_us %>% 
  is.na() %>% 
  colSums()
print(missing) #no missing values

##removing geometry variable (as_tibble) and saving it in csv and RDA format
dat <- svi_us |>
  as_tibble()|>
  select(c(1,3:24)) |>
  rename(geoid10= "geoid")

# save svi data as csv file
write.csv(dat,"svi_us_acs_extract.csv")

# save same file in rda format as backup
save(dat, file = "svi_us_acs.RDA")
```

#PCA Analysis of SVI_US
```{r}
library(pacman)
pacman::p_load(tidyverse, janitor, tidymodels, factoextra, ggfortify, gridExtra, knitr)

dat<- read_csv("svi_us_acs_extract.csv")

svi_pca <- dat[c(5:24)] 

# 
 final <- svi_pca[c(1, 3, 5, 7, 9, 11, 13, 15, 17,19)] #variables for svi_pca analysis
 
# Calculate principal components
#scale: a logical value indicating whether the variables should be scaled to have unit variance before the analysis takes place
pca.ln <- prcomp(final, scale = TRUE)
save(pca.ln, file = "svi_pca_ln.rda")
```

```{r}
#load model object
load("svi_pca_ln.rda")
```

```{r}
eigenvalues_ln <- matrix(pca.ln$sdev^2) #eigenvalues
perc_variance <- round(100*matrix(pca.ln$sdev^2/sum(pca.ln$sdev^2)),1) #variance

#Summary table 
eigenvalues_ln <- cbind(1:10, eigenvalues_ln, perc_variance) 
colum_ln <- c("Principal Component", "Eigenvalues", "Percent Variance")
eigenvalues_ln <- kable(eigenvalues_ln, col.names = colum_ln)
eigenvalues_ln 
```


## Proportion of Variance Plots
The first component explained 81% of the variance 
```{r PVE plots}
#Plots the proportion of variance explained by each component (scree plot)
pve.ln <- pca.ln$sdev^2/sum(pca.ln$sdev^2) #proportion of variance explain by each component
 
#log-transformed data
fviz_eig(pca.ln, main = "Scaled SVI data",
         xlab = "Principal component",
         ylim = c(0,50))

ggsave("svi_pca_screeplot.tiff",
       width=6, height= 6, dpi=300)
```


```{r}
#Extract the first  component with maximum variance
svi_pc1 <- pca.ln$x[, 1]

# Convert to data frame
svi_pc1_df <- data.frame(svi_pc1)

# Save the svi_pc1 as a CSV file
write_csv(svi_pc1_df, "svi_pc1.csv")
```

#combining svi_pc1 with svi data file
```{r}
dat<- read_csv("svi_us_acs_extract.csv") #original svi data
svi_pc1_df <- read_csv("svi_pc1.csv") #SVI PC1

#combining svi_pc1 with svi data file
dat1 <- dat[c(2:22)] # excluding "...1" variable

## final file for svi_pc1 mapping
dat_map<- cbind(dat1, svi_pc1_df)
write_csv(dat_map, "svi_pc1_mapping.csv")

```

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor, tmap, sp, sf, tigris, spdep)
```

#pesticide and pesticide PCA data
```{r}
#join statefp code with dat_fin
state_fips<- read_csv("state_fips.csv")

dat3<- read_csv("mapping_dat_final.csv") |> #pesticide and PCA data adding state fips code
  left_join(state_fips, by="state") |>
  rename(statefp="state_fips_code", countyfp="county_fips_code")
```

#CONUS shapefile to join pesticide/pca data
```{r}
state_list<- list("al", "ar", "az", "ca", "co", "ct", "dc", "de", "fl", "ga", "ia", "id",
                  "il", "in", "ks", "ky", "la", "ma", "md", "me", "mi", "mn", "mo", "ms",
                  "mt", "nc", "nd", "ne", "nh", "nj", "nm", "nv", "ny", "oh", "ok", "or",
                  "pa", "sc", "sd", "tn", "tx", "ut", "va", "vt", "wa", "wi", "wv", "wy", "ri")

us_counties <- counties(state = state_list, cb = FALSE, resolution = "500k", year = "2019", class="sf") |>
  clean_names()

us_counties$countyfp <- as.numeric(us_counties$countyfp)
us_counties$statefp <- as.numeric(us_counties$statefp)
us_counties$geoid <-   as.numeric(us_counties$geoid)
```

```{r}
dat_fin_pest<- left_join(us_counties, dat3, by=c("statefp","countyfp")) |>
  rename(geoid10= "geoid")
```

#Adding statefp code to SVI data
```{r}
svi_fin <- read_csv("svi_pc1_mapping.csv") |>
    left_join(state_fips, by="state") |>
  rename(statefp="state_fips_code")
```

#joining Pesticide PCA data with SVI PCA data
##### Use this dataset for bivariate mapping of pesticide PC1 and Social vulnerability PC1
###########################################################################################################################
```{r}

final_svi_pest <- left_join(dat_fin_pest, svi_fin, by = c("geoid10"))


final_svi_pest_plt<- final_svi_pest |>
  select(c("statefp.x" , "state.y", "countyfp", "countyns", "geoid10", "name",
           "pc_1" , "svi_pc1",  "statefp.y", "geometry"))|>
  mutate(pc_1 = if_else(is.na(pc_1), min(pc_1, na.rm = TRUE), pc_1))|>
  mutate(svi_pc1 = if_else(is.na(svi_pc1), min(svi_pc1, na.rm = TRUE), svi_pc1))

save(final_svi_pest_plt, file="final_svi_pest_bivarate_data.rda")
```


#SVI- PC1 Mapping
```{r}
library(viridis)
library(viridisLite)

svi_index <- "svi_pc1"

# Use a color-blind safe palette from viridis
#color_palette <- viridis_pal()(5)

w <- tm_shape(final_svi_pest_plt) +
  tm_polygons(col = svi_index,
             breaks = quantile(final_svi_pest_plt[[svi_index]], probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = T),
              legend.is.portrait = FALSE,
            palette = viridis::viridis(4, alpha = 1, direction = 1, option = "D"),
              title = "Mapping of SVI Index using ACS 2015-2019 Estimates") +
  tm_borders("gray90", alpha = 0.002) +
  tm_layout(legend.title.size = 1.2,
            legend.text.size = 1) +
  tm_legend(position = c("left", "bottom"), 
            frame = TRUE, stack = "vertical") +
  tm_compass(position = c("right", "top"))  # Add compass to top right corner

tmap_save(w, "US_SVI.tiff",
          width = 12, height = 12, dpi = 150)

```

#Bivariate mapping in ggplot using biscale package
# https://chris-prener.github.io/biscale/articles/biscale.html
```{r}
library(biscale)
library(cowplot)
library(sf)
```

#step 1 preparing data- create bi_class variable 
```{r}
data_test <- bi_class(final_svi_pest_plt, x = pc_1, y = svi_pc1, style = "quantile", dim = 4)
```

#bivariate summary for discussion purposes
```{r}
summary_bi_variate <- data_test |>
  filter(bi_class == "4-4") |>
    as_tibble() |>
  select(c(2,5,6))

table_data <- summary_bi_variate %>%
  group_by(state.y) %>%
  summarise(names = paste(name, collapse = ", "),
            count_names = n()) %>%
  ungroup()

write_csv(table_data, "high_bivariate_val.csv")
```

#step 2- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_test, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 4) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 4,
                    xlab = "Pesticide index ",
                    ylab = "Social vulnerability index",
                    size = 5)+
  theme(text = element_text(size = 6))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("US_bivariate_map.tiff",
       width = 12, height = 12, dpi = 300)
```

# Creating a bivariate map showing only high pesticide and high SVI by setting "bi_class=4-4"
#step 1 preparing data- create bi_class variable 
```{r}
data_test <- bi_class(final_svi_pest_plt, x = pc_1, y = svi_pc1, style = "quantile", dim = 4)

# Filter the data set for rows where bi_class is "4-4"
data_test_filtered <- data_test|>
  filter(bi_class == "4-4") |>
  mutate(geoid = as.character(geoid10))
```

#US county shape file
```{r}
state_list <- list("AL", "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "IA", "ID",
                   "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN", "MO", "MS",
                   "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY", "OH", "OK", "OR",
                   "PA", "SC", "SD", "TN", "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY", "RI")


us_counties <- states(resolution = "500k", 
                     year = "2019", 
                     class="sf") |>
  clean_names() |>
  filter(stusps %in% state_list)
```

#Join filtered counties to shapefile
```{r}
df_map<- st_join(us_counties, data_test_filtered)
```


#step 2- Bivariate Mapping with Biscale
```{r}
# create map
ggplot() +
  geom_sf(data= us_counties, color="black", fill="white", size="0.2")+
  geom_sf(data = data_test_filtered, mapping = aes(fill = bi_class), 
          fill="#CC79A7",
          color = "#CC79A7", 
          size = 0.1, 
          show.legend = FALSE) +
  theme_void()
```



```{r}
ggsave("US_high_high_bivariate_map.tiff",
       width = 12, height = 12, dpi = 300)
```

