---
author: "Jabeen"
title: "Untitled"
output: pdf_document
date: "2023-03-24"
---

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor)
```

# Step 1 Original raw pesticide dataset for 48 US States
```{r}
load("~/Desktop/Jabbeen/jt_dissert/jt_dissert/pest_conus_v2.RDA")
dat_conus$year<- as.numeric(as.character(dat_conus$year))
#write_csv(dat, "dat_pest_conusv1.csv")
```

# Step 2 creating variables median, temporal and spatial variables using original dataset 03/30/23
```{r}
#median quantity and temporal coverage of chemicals applied
median_pest <- dat_conus %>%
  group_by(state_name, county_fips_code, compound) %>%
  summarize(median_pest = median(pest_applied, na.rm = TRUE),
            n_yrs_repeat = sum(!is.na(pest_applied)),
            years = list(unique(year[!is.na(pest_applied)])))|>
  mutate(pct_yrs= round((n_yrs_repeat/28)*100, 2))


#spatial coverage using states (percent states with any compound applied per year)
state_cover <- median_pest %>%
  group_by(compound) %>%
  summarize(state_cvr = n_distinct(state_name[!is.na(median_pest)])) |>
  mutate(pct_state_cvr = round((state_cvr/48)*100, 2))

#join meadian_pest and state-cover
dat_med_st_cvr<- left_join(median_pest, state_cover, by=c("compound"))

######################################################################
#Spatial coverage of chemicals applied (coverage that shows percent counties)
summ <- median_pest %>%
    group_by(state_name, compound) %>%
    summarize(n_counties_pest = ifelse(any(!is.na(median_pest)),
                                       n_distinct(county_fips_code[!is.na(median_pest)]), 0))

#number of counties
n_cnty<- median_pest %>%
  group_by(state_name) %>%
  summarize(num_counties = n_distinct(county_fips_code))

#join n_cnty with spatial coverage to get percent spatial coverage
summ_spat_pct<- left_join(summ,n_cnty, by="state_name") |>
  mutate(pct_spat_cover = round((n_counties_pest/num_counties)*100, 2))
  

#########################################################################

# Final data set that has median pesticide applied, pct spatial and temporal coverage

chem_appl<- left_join(dat_med_st_cvr, summ_spat_pct, 
                      by=c("state_name","compound"))

chem_appl<- chem_appl |>
  rename(state=state_name, pct_state_cvr = pct_state_cvr,
         pct_cnty_cover=pct_spat_cover) |>
  mutate(years = gsub("(^c\\(|\\)$)", "", years))

# save same file in rda format as backup
save(chem_appl, file = "chem_appl_spat_temp_cvr_v3.RDA")
```

#Step 3 chemical selection function
```{r}
compound_filter <- function(data, pct_yrs_thresh, pct_state_cvr_thresh, pct_cnty_cvr_thresh) {
  
  # Filter the data based on pct_yrs threshold
  data <- data %>% group_by(compound) %>% filter(mean(pct_yrs) >= pct_yrs_thresh)
  
  # Filter the data based on pct_state_cvr threshold
  data <- data %>% group_by(compound) %>% filter(mean(pct_state_cvr) >= pct_state_cvr_thresh)
  
  # Filter the data based on pct_cnty_cvr threshold
  data <- data %>% group_by(compound) %>% filter(mean(pct_cnty_cover) >= pct_cnty_cvr_thresh)
  
  # Return unique list of compounds that match the criteria as a dataframe
  unique(data$compound) %>% as.data.frame()
}
```

# Different percentages generate different lists
```{r}
#Use file: load("chem_appl_spat_temp_cvr_v3.RDA")
compound_filter<- compound_filter(chem_appl, 80, 80, 80)
compound_filter<- compound_filter |> rename(compound = ".")
write_csv(compound_filter, "chemical_list80_80_80.csv")
```


# Step 4 Creating a final subset of dataset with selected chemicals 05/03/23
```{r}
# read in the dataset
load("chem_appl_spat_temp_cvr_v3.RDA")

# Filter chem_appl based on the values in compound_filter
chem_appl_filtered <- chem_appl %>%
  filter(compound %in% compound_filter$compound)

# Summarize pct_yrs by compound
summary <- chem_appl_filtered %>%
  group_by(compound) %>%
  summarize(
    mean_pct_yrs = mean(pct_cnty_cover),
    median_pct_yrs = median(pct_cnty_cover),
    max_pct_yrs = max(pct_cnty_cover),
    min_pct_yrs = min(pct_cnty_cover),
    n = n()
  )

save(chem_appl_filtered, file="chem_appl_filtered_v4.RDA")
```

# Step 5 Grouping the pesticides by class and creating final dataset
```{r}

# create a named vector of compound classes
compound_classes <-list(Herbicide = c("2,4-D", "ATRAZINE", "ACETOCHLOR","CLETHODIM", "DICAMBA","GLYPHOSATE", 
                                      "IMAZETHAPYR", "PARAQUAT", "PENDIMETHALIN","CHLORIMURON","METOLACHLOR",
                                      "METRIBUZIN","METSULFURON", "NICOSULFURON", "PICLORAM", "SETHOXYDIM", 
                                      "SIMAZINE","THIFENSULFURON","TRIBENURON METHYL", "TRICLOPYR", "TRIFLURALIN"),
                        
                        Insecticide = c("CARBARYL", "CHLORPYRIFOS", "CYHALOTHRIN-LAMBDA",
                                        "DIMETHOATE", "ESFENVALERATE","PERMETHRIN"),
                        
                        Fungicide = c("CHLOROTHALONIL","MANCOZEB","PROPICONAZOLE"))

# Create a new variable "class" in chem_appl_filtered
chemi_class <- chem_appl_filtered %>% 
  mutate(class = case_when(
    compound %in% compound_classes[["Herbicide"]] ~ "Herbicide",
    compound %in% compound_classes[["Insecticide"]] ~ "Insecticide",
    compound %in% compound_classes[["Fungicide"]] ~ "Fungicide",
    TRUE ~ NA_character_
  ))

# Convert "class" variable to factor
chemi_class$class <- as.factor(chemi_class$class)

# Print the first few rows of the resulting dataset
head(chemi_class)

write_csv(chemi_class, "final_chemv5.csv")

# save same file in rda format as backup
save(chemi_class, file = "final_chem_v5_backup.RDA")
```


06/05/24 # For summary characteristics of pesticides table
```{r}
library(dplyr)
dat <- load("final_chem_v5_backup.RDA")

# Assuming 'chemi_class' is dataframe
chemi_class_grouped <- chemi_class %>%
  group_by(compound) %>%
  summarize(
    median = mean(median_pest, na.rm = TRUE), # Calculate the median by taking the average of median
    class = first(class), # Get the class for each compound
    .groups = 'drop' # Drop grouping information
  ) %>%
  mutate(
    median = round(median, 2), # Round median_pest to two decimal places
    rank = rank(-median) # Create the rank based on the rounded median
  ) %>%
  select(rank, compound, class, median) %>%
  arrange(rank)

# Print the data to see the result
print(chemi_class_grouped, n = 30)

# save as csv format

write_csv(chemi_class_grouped, "aim1_us_final_chem_median_mean_char.csv")


# Approach 2, taking the sum of median instead of mean

# Assuming 'chemi_class' is dataframe
chemi_grouped <- chemi_class %>%
  group_by(compound) %>%
  summarize(
    amount = sum(median_pest, na.rm = TRUE), # Calculate the amount by taking the sum of median
    class = first(class), # Get the class for each compound
    .groups = 'drop' # Drop grouping information
  ) %>%
  mutate(
    amount = round(amount, 2), # Round median_pest to two decimal places
    rank = rank(-amount) # Create the rank based on the rounded median
  ) %>%
  select(rank, compound, class, amount) %>%
  arrange(rank)

# Print the data to see the result
print(chemi_grouped, n = 30)

# save as csv format

write_csv(chemi_grouped, "aim1_us_final_chem_median_sum_char.csv")




```

