---
title: "02.2_nebraska_pest_part2"
author: "Taiba"
date: "2023-06-17"
output: html_document
---
############################################################################################################################################################################################ Don't use these part 2 chemicals for analysis/use only part 1 (90%, 100%)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor,reshape2)
```

# 1 Loading original conus data 1992-2019
```{r}

load("~/Desktop/Jabbeen/jt_dissert/jt_dissert/pest_conus_v2.RDA")
dat_conus$year<- as.numeric(as.character(dat_conus$year))
```

# Step 1 Filtering pesticides for NE from 1992- 2014 has 205956 obs and 6 variables
```{r}
dat_ne <- dat_conus |>
  filter(state_fips_code == "31" & year %in% c(1992, 1993, 1994, 1995, 1996, 1997, 1998, 1999, 2000,
                                               2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009,
                                               2010, 2011, 2012, 2013, 2014))

  #save(dat_ne, file = "pest_ne_v1.RDA")
```

# Step 2 creating variables median, temporal and spatial variables using original dataset 03/30/23
```{r}
load("pest_ne_v1.RDA")
#median quantity and temporal coverage of chemicals applied
median_pest <- dat_ne %>%
  group_by(county_fips_code, compound) %>%
  summarize(median_pest = median(pest_applied, na.rm = TRUE),
            n_yrs_repeat = sum(!is.na(pest_applied)),
            years = list(unique(year[!is.na(pest_applied)])))|>
  mutate(pct_yrs= round((n_yrs_repeat/23)*100, 2))


#spatial coverage using counties (percent counties with any compound applied per year)
county_cover <- median_pest %>%
  group_by(compound) %>%
  summarize(county_cvr = n_distinct(county_fips_code[!is.na(median_pest)])) |>
  mutate(pct_county_cvr = round((county_cvr/93)*100, 2))

#join median_pest and county-cover
dat_med_cnty_cvr<- left_join(median_pest, county_cover, by=c("compound"))

#########################################################################

# Final data set that has median pesticide applied, pct spatial and temporal coverage has 18730 obs and 8 variables

chem_appl<- dat_med_cnty_cvr |>
  mutate(years = gsub("(^c\\(|\\)$)", "", years))

# save same file in rda format as backup
#save(chem_appl, file = "ne_chem_appl_spat_temp_cvr_v2.RDA")
```

#Step 3 chemical selection function
```{r}
compound_filter <- function(data, pct_yrs_thresh, pct_cnty_cvr_thresh) {
  
  # Filter the data based on pct_yrs threshold
  data <- data %>% group_by(compound) %>% filter(mean(pct_yrs) >= pct_yrs_thresh)
  
  # Filter the data based on pct_cnty_cvr threshold
  data <- data %>% group_by(compound) %>% filter(mean(pct_county_cvr) >= pct_cnty_cvr_thresh)
  
  # Return unique list of compounds that match the criteria as a dataframe
  unique(data$compound) %>% as.data.frame()
}
```

# Different percentages generate different lists
```{r}
#Use file: load("ne_chem_appl_spat_temp_cvr_v2.RDA")
compound_filter<- compound_filter(chem_appl,80,80)
compound_filter<- compound_filter |> rename(compound = ".")

write_csv(compound_filter, "ne_chemical_list80_80.csv")

```


# Step 4 Creating a final subset of dataset with selected chemicals 80-80 and saving as part 2 list (part 1 in 02.1_nebraska_pest_analysis.Rmd was 90-100)
```{r}
# read in the dataset
load("ne_chem_appl_spat_temp_cvr_v2.RDA")

# Filter chem_appl based on the values in compound_filter
chem_appl_filtered <- chem_appl %>%
  filter(compound %in% compound_filter$compound)

# Summarize pct_yrs by compound
summary <- chem_appl_filtered %>%
  group_by(compound) %>%
  summarize(
    mean_pct_yrs = mean(pct_county_cvr),
    median_pct_yrs = median(pct_county_cvr),
    max_pct_yrs = max(pct_county_cvr),
    min_pct_yrs = min(pct_county_cvr),
    n = n()
  )

save(chem_appl_filtered, file="ne_chem_appl_filtered_part2.RDA")
```

# Step 5 Grouping the pesticides by class and creating final dataset has 2976 obs and 9 variables
```{r}

# create a named vector of compound classes
compound_classes <-list(Herbicide = c("2,4-D","2,4-DB", "ATRAZINE","ALACHLOR","ACETOCHLOR","ACIFLUORFEN","BROMOXYNIL","BENTAZONE", 
"CHLORIMURON","CHLORSULFURON", "CLETHODIM","CLOPYRALID", "DICAMBA","DIMETHENAMID", "EPTC","FLUAZIFOP","FLUMETSULAM","FOMESAFEN", "GLYPHOSATE", "HALOSULFURON", "IMAZAQUIN", 
                                      "IMAZETHAPYR","LACTOFEN", "MCPA","METOLACHLOR","PARAQUAT", "METRIBUZIN","METSULFURON", "NICOSULFURON", "PICLORAM","PENDIMETHALIN","PRIMISULFURON", "QUIZALOFOP", "RIMSULFURON", "SETHOXYDIM", "THIFENSULFURON","TRIASULFURON", "TRIFLURALIN", "TRIBENURON METHYL"),
                        
                        Insecticide = c( "BIFENTHRIN" ,"CHLORPYRIFOS", "CYFLUTHRIN","CYHALOTHRIN-LAMBDA", "DIMETHOATE", "ESFENVALERATE","PERMETHRIN","TEFLUTHRIN","TERBUFOS"),

Fungicide = c("PROPICONAZOLE"))

# Create a new variable "class" in chem_appl_filtered
chemi_class <- chem_appl_filtered %>% 
  mutate(class = case_when(
    compound %in% compound_classes[["Herbicide"]] ~ "Herbicide",
    compound %in% compound_classes[["Insecticide"]] ~ "Insecticide",
    compound %in% compound_classes[["Fungicide"]] ~ "Fungicide",
    TRUE ~ NA_character_
  ))

# Convert "class" variable to factor
chemi_class$class <- as.factor(chemi_class$class)

# Print the first few rows of the resulting dataset
head(chemi_class)

write_csv(chemi_class, "ne_final_chem_part2.csv")

# save same file in rda format as backup
save(chemi_class, file = "ne_final_chem_part2.RDA")

```


# Step 6 boxplots for each chemical
```{r}
load("ne_final_chem_part2.RDA")

ggplot(chemi_class, aes(y = compound, x = log(median_pest))) +
  geom_boxplot() +
  labs(title = "",
       x = "Median pesticide applied (1992 - 2014) - Log scale", 
       y = "") +
  facet_grid(factor(class, levels=c("Herbicide", "Insecticide", "Fungicide"))~ ., scales = "free_y", 
             switch = "y", space = "free_y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, size = 10.5),
        axis.text.y = element_text(size = 10.5),
        strip.text = element_text(colour = "black", size = 10.5),
        strip.text.x = element_text(colour = "black", size = 10.5),
        strip.text.y = element_text(colour = "black", size = 7))


ggsave("part2_ne_chemical_box_plt.tiff", 
       width = 6,height = 9,
       dpi=300)
```

#Step 7 correlation matrix
```{r}
# pivot from long to wide format
chemi_class_wide <- chemi_class[c(1:3)] %>%
  pivot_wider(names_from = compound, values_from = median_pest)|>
   mutate_at(vars(2:33), ~log10(.+0.0000001)) 

write_csv(chemi_class_wide, "ne_final_chem_wide_part2.csv")

# save same file in rda format as backup
save(chemi_class_wide, file = "ne_final_chem_wide_part2_backup.RDA")
```


```{r}
#correlation matrix
cormat<- round(x=cor(chemi_class_wide[c(2:33)], method = "spearman", 
                     use= "complete.obs"), digits=2) |>
  melt()


ggplot(cormat, aes(x=Var2, y=Var1, fill=value)) +
  geom_tile(color="white")+
  scale_fill_gradient2(low="red", high="blue", mid="white",
                       midpoint=0,
                       limit=c(-1,1), space= "Lab",
                       name="Spearman Correlation | Pesticides applied in NE Counties (1992-2014)")+
  geom_text(aes(label = round(value, 2)), size = 3, color = "black") + # Add text labels
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
        axis.text.y = element_text(angle = 0, vjust = 1, size = 12, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom", legend.box = "horizontal")+
  coord_fixed()

ggsave("part2ne_chemical_correlations.tiff",
       width=12, height= 10, dpi=300)
```