---
title: "05_final_data_analysis"
author: "Taiba"
date: "2023-07-23"
output: pdf_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse, dplyr, janitor,gWQS)
```

#sdoh data
```{r}
#join variables: geoid10 and county name
load("ped_sdohProcessed.RDA") |> as_tibble()

sdoh <- dat |>
  as_tibble() |>
  rename(geoid10= "geoid") |>
  select(-c(3:20)) |>
  mutate_at(vars(3:11), ~ round(., 2)) |>
  mutate(name = str_split_fixed(name, ", ", 2)[, 1])|>
  rename(county= "name")

rm(dat)
```

#pediatric population
```{r}
ped_pln<- read_csv("ped_popln_exp.csv") 
```

#pediatric cancer data
```{r}
ped_can_count<- read_csv("ped_cancer_v3.csv") |>
  select(c(1:13))
```

#birth defects dadta
```{r}
#join variable: county name
bd_count<- read_csv("bd_cnty_v7.csv") 
```

#join sdoh with outcomes
```{r}
sdoh_outcms <- sdoh |>
  full_join(ped_can_count, by = "county") |>
  full_join(ped_pln, by = "county") |>
  full_join(bd_count, by = "county") |>
  clean_names()

```

#chemical data
```{r}
ne_chem<- read_csv("ne_final_chem_widev5.csv") |>
  rename(geoid10= "county_fips_code") |>
  mutate(geoid10 = str_pad(as.character(geoid10), width = 3, pad = "0"),
         geoid10 = paste0("31", geoid10))

```


# join chemicals, sdoh (percentage variables) and outcomes into a file
```{r}
fin_dat<- full_join(ne_chem, sdoh_outcms, by= "geoid10") |>
  select(-c("geoid")) |>
  mutate_at(vars(44:72), ~ replace_na(., 0))

write_csv(fin_dat, "final_dissrt_analy_data.csv")
```

# join sdoh count data to the final_dissrt_analy_data for stratified analysis purpose

```{r}
dat1 <- read_csv("final_dissrt_analy_data.csv")
dat2 <- read_csv("ped_sdohcount.csv")
 
# 
dat2 <- dat2 |>
  select(-contains("...1"))

#
fin_dat1 <- full_join(dat1, dat2, by= "geoid10") |>
  mutate_at(vars(73:82), ~ replace_na(., 0))

write_csv(fin_dat1, "final_dissrt_analy_sdh_count_datav1.csv")

 # Save same file as RDA 
save(fin_dat1, file = "final_dissrt_analy_sdh_count_datav1.RDA")
```

#############################################################################
```{r}
pacman::p_load(tidyverse, janitor, gWQS)
```

# using data without sdoh counts
```{r}
dat<- read_csv("final_dissrt_analy_data.csv")
```

#model
```{r, warning=FALSE}
chem_mix<- names(dat)[2:33]

wqs_res<- gwqs(ophthal ~ wqs+ m_blk_pct + f_blk_pct + m_hisp_pct + f_hisp_pct + m_heal_ins_pct + 
                 f_heal_ins_pct +vehicle_pct + sing_pent_pct + ssi_snap_pct, 
               mix_name = chem_mix, 
               offset= dat$live_birth,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = TRUE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

#gwqs_summary_tab(wqs_res)
save(wqs_res, file="~/Dissertation/test/birth_def/nrh_ophthal_bd.RDA")
```


```{r}
gwqs_scatterplot(wqs_res)
gwqs_summary_tab(wqs_res)

gwqs_barplot(wqs_res)
gwqs_weights_tab(wqs_res)
```

#repeated holdout # try to use rs- subset term
```{r, warning=FALSE}
wqs_res_rh <- gwqsrh(cardiac ~ wqs+
                       #Below are list of covariates included
                       m_blk_pct + f_blk_pct + m_hisp_pct + f_hisp_pct + m_heal_ins_pct +
                       f_heal_ins_pct +vehicle_pct + sing_pent_pct + ssi_snap_pct, 
                     ###################
                     mix_name = chem_mix, #list of pesticides as mixture
                     ###################
                     offset=dat$live_birth, #offset variable to yield rat
                     ###################
                     zero_infl = FALSE, #only for sub-types
                     ###################
                     data = dat, #input dataset
                     ###################
                     q=10, #Converting pesticides to quantile scale
                     ###################
                     validation= 0.6, #split 60% data for validation
                     ###################
                     b = 10, #considering 100 bootstrap samples
                     ###################
                     b1_pos = TRUE, #assuming beta as positive
                     ###################
                     b1_constr = FALSE, #Unconstrained beta direction
                     ###################
                     family= "negbin", # assuming outcome following negative binomial distribution
                     ###################
                     seed = 2023, #seed to reproduce the same test and training data
                     ###################
                     rh=5 #repeated holdouts
                     ###################
                     )
```

```{r}
gwqs_summary_tab(wqs_res_rh)
gwqsrh_boxplot(wqs_res_rh, tau = 0)
gwqs_weights_tab(wqs_res_rh)

#ggsave("gwqsrh_abdominal_bd.tiff", 
       #width = 6,height = 9,
       #dpi=300)
```


```{r}
gwqs_loop <- function(input_df, output_folder, outcomes, covariates=NULL, 
                      chem_mix) {
  # Loop through the list of outcomes
  for (outcome in outcomes) {
    # Construct the formula with specified covariates
    formula <- paste(outcome, "~ wqs ", covariates)
    
    # Create the file names based on the outcome and model type
    filename_prefix <- "model_output"
    file_name_wqs <- paste0("nrh", "_", outcome, ".rda")
    file_name_wqs_rh <- paste0("rh", "_", outcome, ".rda")
    
    # Perform the first model (gWQS)
    wqs_res <- tryCatch(
      gwqs(as.formula(formula), mix_name = chem_mix, offset = dat$ped_popln,
           zero_infl = FALSE, data = input_df, q = 10, validation = 0.6, b = 10,
           b1_pos = TRUE, b1_constr = FALSE, family = "negbin", seed = 2023),
      error = function(e) {
        cat(paste("Error occurred for outcome:", outcome, "\n", "Error message:", e$message, "\n"))
        return(NULL)
      }
    )
    
    if (is.null(wqs_res)) {
      next  # Move to the next outcome variable
    } else {
      # Save the first model output to RDA file
      save(wqs_res, file = file.path(output_folder, file_name_wqs))
    }
    
    # Perform the second model (gWQSRH)
    wqs_res_rh <- tryCatch(
      gwqsrh(as.formula(formula), mix_name = chem_mix, offset = dat$ped_popln,
             zero_infl = FALSE, data = input_df, q = 10, validation = 0.6, b = 100,
             b1_pos = TRUE, b1_constr = FALSE, family = "negbin", seed = 2023, rh = 5),
      error = function(e) {
        cat(paste("Error occurred for outcome:", outcome, "\n", "Error message:", e$message, "\n"))
        return(NULL)
      }
    )
    
    if (is.null(wqs_res_rh)) {
      next  # Move to the next outcome variable
    } else {
      # Save the second model output to RDA file
      save(wqs_res_rh, file = file.path(output_folder, file_name_wqs_rh))
    }
  }
}


```

#run function
```{r, warning=FALSE}

gwqs_loop(input_df = dat,
          output_folder = "~/Dissertation/test", 
          outcomes = names(dat)[44:55],  #update location of outcomes
          covariates = "+ m_blk_pct + f_blk_pct + m_hisp_pct + f_hisp_pct + m_heal_ins_pct + f_heal_ins_pct +vehicle_pct + sing_pent_pct + ssi_snap_pct ",
          chem_mix = names(dat)[2:33])
```

```{r}
load("~/Dissertation/test/nrh_total_can.rda") 

summary(wqs_res)
gwqs_barplot(wqs_res)

```


```{r}
load("~/Dissertation/test/rh_leukemia.rda") 

summary(wqs_res_rh)
gwqsrh_boxplot(wqs_res_rh)
gwqs_weights_tab(wqs_res_rh)

#
ggsave("gwqsrh_leukemia.tiff", 
       width = 6,height = 9,
       dpi=300)
```


```{r}
load("~/Dissertation/test/birth_def/nrh_overall_bd.rda") 
load("~/Dissertation/test/birth_def/nrh_cardiac.rda") 
load("~/Dissertation/test/birth_def/nrh_cranial_bd.rda")
load("~/Dissertation/test/birth_def/nrh_hepatic_bd.rda") 
load("~/Dissertation/test/birth_def/nrh_musc_ske.rda")
load("~/Dissertation/test/birth_def/nrh_ophthal_bd.rda")
load("~/Dissertation/test/birth_def/nrh_renal_genuri.rda")


total_bd<- wqs_res$final_weights |>
  mutate(outcome = "total_bd")
rm(wqs_res)

cardiac<- wqs_res$final_weights |>
  mutate(outcome = "cardiac")
rm(wqs_res)

cranial<- wqs_res$final_weights |>
  mutate(outcome = "cranial")
rm(wqs_res)

hepatic<- wqs_res$final_weights |>
  mutate(outcome = "hepatic")
rm(wqs_res)

musc_ske<- wqs_res$final_weights |>
  mutate(outcome = "musc_ske")
rm(wqs_res)

ophthal<- wqs_res$final_weights |>
  mutate(outcome = "ophthal")
rm(wqs_res)
  
renal_genuri<- wqs_res$final_weights |>
  mutate(outcome = "renal_genuri")
rm(wqs_res)

# Get a list of all dataframes in the global environment
dfs <- mget(ls())

# Combine all dataframes into a single dataframe
combined_df <- bind_rows(dfs)|>
  clean_names()|>
  mutate(mix_name = str_to_title(mix_name),
         mean_weight = sprintf("%.2f", mean_weight)) |>
  #select(-c(2:4)) |>
  pivot_wider(names_from = outcome, values_from = mean_weight) |>
  select(mix_name, total_bd, cranial, cardiac, renal_genuri, musc_ske, hepatic,ophthal)

write_csv(combined_df, "~/Dissertation/test/birth_def/bd_weights.csv")
```




```{r}
load("~/Dissertation/test/unadj/nrh_benign_neop_mass.rda") 

summary(wqs_res)

```

# Figures for aim2 ped cancer- WQS estimates 01/07/24
```{r}
dat <- read_csv("ped_can_wqs.csv") |>
  clean_names() |>
  mutate(cancer_type = factor(cancer_type, levels = c("Total cancer", "Brain and other CNS", 
                                                        "Leukemia", "Malignant bone", "Germ cell")))|>
  mutate(sign = if_else(lower_limit <= 0 & upper_limit >= 0, "non_significant", "significant"))|>
  rename(`Statistical significance` = sign)
  

```

```{r}
cbbPalette <- c("#e34234", "royalblue")

ggplot(dat, aes(x = as.factor(cancer_type), y = estimates, ymin = lower_limit, 
                ymax = upper_limit, linetype=model)) + 
  geom_pointrange(aes(col = `Statistical significance`), position=position_dodge(width=0.75),size = 0.5) +
  ylab(expression(paste(beta[WQS]," - Chemical mixture [95% CI]"))) +
  scale_y_continuous(breaks = seq(-0.2, 1.0, 0.2),limits=c(-0.2, 1.0))+
  geom_hline(aes(yintercept = 0), linetype="dashed", color="gray60") + 
  scale_colour_manual(values=cbbPalette) + 
  ggtitle("")+
  xlab("")+
  theme_bw()+
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust = 0.5),
        text=element_text(size=12,  family="Arial Black"),
        axis.text.y = element_text(size = 12, family="Arial Black"),
        axis.title.y = element_text(size = 12, family = "Arial Black"),
        legend.text.align = 0,
        strip.text=element_text(size=12, face="bold",family="Arial Black"),
        strip.placement = "outside",
        axis.ticks.x = element_blank(),
        legend.position = "bottom")+
  guides(color=guide_legend(override.aes = list(shape=15, size=1.5)))
```

```{r}
ggsave("pead_can_wqs.tiff",width = 8,height = 4)
```

# creating plot for constrained and unconstrained model
```{r}
df<- read.csv("aim2_main_result.csv") |>
  mutate(sign = if_else(ci_ll <= 0 & ci_ul >= 0, "non_significant", "significant"))|>
  rename(`Statistical significance` = sign) |>
  mutate(cancer = factor(cancer, levels = c("Total Cancer", "Brain and Other CNS", 
                                                        "Leukemia", "Malignant Bone", "Germ Cell")))


cbbPalette <- c("#e34234", "royalblue")

ggplot(df, aes(x = as.factor(cancer), 
               y = beta, ymin = ci_ll,  ymax = ci_ul, 
               shape = model_type)) + 
  geom_pointrange(aes(col = `Statistical significance`), 
                  position=position_dodge(width=0.75),size = 0.5) +
  ylab(expression(paste(beta[WQS]," - Chemical mixture [95% CI]"))) +
  scale_y_continuous(breaks = seq(-0.2, 0.8, 0.2),limits=c(-0.2, 0.9))+
  geom_hline(aes(yintercept = 0), linetype="dashed", color="gray60") + 
  scale_colour_manual(values=cbbPalette) + 
  ggtitle("")+
  xlab("")+
  labs(shape="gWQS model")+
  theme_bw()+
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust = 0.5),
        text=element_text(size=12,  family="Arial Black"),
        axis.text.y = element_text(size = 12, family="Arial Black"),
        axis.title.y = element_text(size = 12, family = "Arial Black"),
        legend.text.align = 0,
        strip.text=element_text(size=12, face="bold",family="Arial Black"),
        strip.placement = "outside",
        axis.ticks.x = element_blank(),
        legend.position = "bottom")+
  guides(color=guide_legend(override.aes = list(shape=15, size=1.5)))+
  guides(shape = guide_legend(ncol = 2), color = guide_legend(ncol = 2))
```

```{r}
ggsave("pead_can_final_wqs.tiff",width = 10,height = 4, dpi=300)
```

########################################################################################################################################################################################################
# Aim3 results tables-did not create plots
# gWQS repeated holdout with constrains
```{r}
gwqs_loop <- function(input_df, output_folder, outcomes, covariates=NULL, 
                      chem_mix) {
  # Loop through the list of outcomes
  for (outcome in outcomes) {
    # Construct the formula with specified covariates
    formula <- paste(outcome, "~ wqs ", covariates)
    
    # Create the file names based on the outcome and model type
    filename_prefix <- "model_output"
    file_name_wqs <- paste0("nrh", "_", outcome, ".rda")
    file_name_wqs_rh <- paste0("rh", "_", outcome, ".rda")
    
    # Perform the first model (gWQS)
    wqs_res <- tryCatch(
      gwqs(as.formula(formula), mix_name = chem_mix, offset = dat$live_birth,
           zero_infl = FALSE, data = input_df, q = 10, validation = 0.6, b = 10,
           b1_pos = TRUE, b1_constr = TRUE, family = "negbin", seed = 2023),
      error = function(e) {
        cat(paste("Error occurred for outcome:", outcome, "\n", "Error message:", e$message, "\n"))
        return(NULL)
      }
    )
    
    if (is.null(wqs_res)) {
      next  # Move to the next outcome variable
    } else {
      # Save the first model output to RDA file
      save(wqs_res, file = file.path(output_folder, file_name_wqs))
    }
    
    # Perform the second model (gWQSRH)
    wqs_res_rh <- tryCatch(
      gwqsrh(as.formula(formula), mix_name = chem_mix, offset = dat$live_birth,
             zero_infl = FALSE, data = input_df, q = 10, validation = 0.6, b = 100,
             b1_pos = TRUE, b1_constr = TRUE, family = "negbin", seed = 2023, rh = 5),
      error = function(e) {
        cat(paste("Error occurred for outcome:", outcome, "\n", "Error message:", e$message, "\n"))
        return(NULL)
      }
    )
    
    if (is.null(wqs_res_rh)) {
      next  # Move to the next outcome variable
    } else {
      # Save the second model output to RDA file
      save(wqs_res_rh, file = file.path(output_folder, file_name_wqs_rh))
    }
  }
}


```

#run function
```{r, warning=FALSE}

gwqs_loop(input_df = dat,
          output_folder = "~/Dissertation/test/rh_bd_constrain", 
          outcomes = names(dat)[58:72],  #update location of outcomes
          covariates = "+ m_blk_pct + f_blk_pct + m_hisp_pct + f_hisp_pct + m_heal_ins_pct + f_heal_ins_pct +vehicle_pct + sing_pent_pct + ssi_snap_pct ",
          chem_mix = names(dat)[2:33])
```

```{r}
load("~/Dissertation/test/nrh_cardiac.rda") 

summary(wqs_res)
gwqs_barplot(wqs_res)

```

# Creating a table for 
```{r}
load("~/Dissertation/test/cancer/rh_total_can.rda") 
load("~/Dissertation/test/cancer/rh_leukemia.rda") 
load("~/Dissertation/test/cancer/rh_cns_and_miscellaneous_intracranial_and_intraspinal.rda")
load("~/Dissertation/test/cancer/rh_malignant_bone_tumors.rda") 
load("~/Dissertation/test/cancer/rh_germ_cell_tumors_trophoblastic_tumors_and_neopla.rda")


total_can<- wqs_res_rh$final_weights |>
  mutate(outcome = "total_cancer")
rm(wqs_res_rh)

leukemia<- wqs_res_rh$final_weights |>
  mutate(outcome = "leukemia")
rm(wqs_res_rh)

cns<- wqs_res_rh$final_weights |>
  mutate(outcome = "cns")
rm(wqs_res_rh)

malg_bone<- wqs_res_rh$final_weights |>
  mutate(outcome = "malg_bone")
rm(wqs_res_rh)

germ_cell<- wqs_res_rh$final_weights |>
  mutate(outcome = "germ_cell")
rm(wqs_res_rh)
  

# Get a list of all dataframes in the global environment
dfs <- mget(ls())

# Combine all dataframes into a single dataframe
combined_df <- bind_rows(dfs)|>
  clean_names()|>
  mutate(mix_name = str_to_title(mix_name),
         across(where(is.numeric), ~as.numeric(sprintf("%.2f", .))),
         summary = sprintf("%.2f (%.2f, %.2f)", estimate, x2_5_percent, x97_5_percent)) |>
  select(-c(2:4)) |>
  pivot_wider(names_from = outcome, values_from = summary) |>
  select(mix_name, total_cancer, cns, leukemia, malg_bone, germ_cell)

write_csv(combined_df, "cancer_gwqs_rh_weights.csv")
```


# creating plot for constrained and unconstrained model/non-repeated holdout
```{r}
df <- read.csv("aim3_main_results.csv") |>
  mutate(sign = if_else(p_val <= 0.05, "significant", "non_significant")) |>
  rename(`Statistical significance` = sign) |>
  mutate(Birth_defect = factor(Birth_defect, levels = c("Total defects", "Cranial", 
                                                        "Cardiac", "Renal", "Musculoskeletal", "Hepatic", "Ophthalmic"))) |>
  mutate(ymin = beta - 1.96 * `std_err`, ymax = beta + 1.96 * `std_err`)

cbbPalette <- c("royalblue", "#e34234")

ggplot(df, aes(x = as.factor(Birth_defect), y = beta, ymin = ymin, ymax = ymax, shape = model_type)) + 
  geom_pointrange(aes(color = `Statistical significance`), position = position_dodge(width = 0.75), size = 0.5) +
  ylab(expression(paste(beta[WQS], " - Chemical mixture [Std Err]"))) +
  scale_y_continuous(breaks = seq(-0.2, 0.8, 0.2), limits = c(-0.2, 0.9)) +
  geom_hline(aes(yintercept = 0), linetype = "dashed", color = "gray60") + 
  scale_colour_manual(values = cbbPalette) + 
  ggtitle("") +
  xlab("") +
  labs(shape = "gWQS model") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust = 0.5),
        text = element_text(size = 10, family = "Arial Black"),
        axis.text.y = element_text(size = 10, family = "Arial Black"),
        axis.title.y = element_text(size = 10, family = "Arial Black"),
        legend.text.align = 0,
        strip.text = element_text(size = 10, face = "bold", family = "Arial Black"),
        strip.placement = "outside",
        axis.ticks.x = element_blank(),
        legend.position = "bottom") +
  guides(color = guide_legend(override.aes = list(shape = 15, size = 1.5)),
         shape = guide_legend(ncol = 2), 
         color = guide_legend(ncol = 2))

```

```{r}
ggsave("birth_def_final_wqs.tiff",width = 10,height = 4, dpi=300)
```


