---
title: "07.WQS_mixture_pediatric_cancer072923"
author: "Taiba"
date: "2023-07-29"
output: pdf_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse,broom, tidyr, dplyr, janitor,gWQS)
```

#Load final dissertation data with sdoh count and percentages
```{r}
dat<- read_csv("final_dissrt_analy_sdh_count_datav1.csv")
```

# define chemicals to include in the mixture
#correlation and log transformation already done 
```{r}
chem_mix<- names(dat)[2:33] # dat for log-transformed scale
# dataquant for decile chemical scale data
```

# Quantile-transformation of chemical mixture
```{r}
dataquant <- dat |>
  mutate(across(all_of(chem_mix), ~ntile(., 10))) |>
  rename_with(~paste0(.x, "_quantile"), all_of(chem_mix))

```

## llinear model w/ one chemical- guassian does not work for my data (dataset is for poisson model)

## multiple linear regression- guassian does not work for my data (dataset is for poisson model)

```{r}
chemicals
multipleGLM = glm(total_can ~ m_blk_pct + f_blk_pct + m_hisp_pct + f_hisp_pct + m_heal_ins_pct + f_heal_ins_pct + sing_pent_pct + ssi_snap_pct +  vehicle_pct, data = dat)
summary(multipleGLM)
matrix_coef <- summary(multipleGLM)$coefficients  # Extract coefficients in matrix
matrix_coef   

#write_csv(matrix_coef, "totalcan_multiplereg.csv")
```
############################################################################################
###############################################################################################
#  Analysis strategy 1
#WQS regression unadjusted without covariates
# fit an unadjusted model to look at the association between the chemicals and the outcome
# outcome = outcome variable/cancer/birth defects
## there's a signal only in the positive direction
#summary(dat[,chemicals])
```{r}
pacman::p_load(tidyverse, janitor, gWQS)

chem_mix<- names(dat)[2:33]
offset <- dat$live_birth

results1 = gwqs(total_bd ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$live_birth,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = TRUE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 > 0) 

summary(results1$fit)
results1$final_weights
```

## there's a signal only in the positive direction/ negative non-significant
```{r}
chem_mix<- names(dat)[2:33]
offset <- dat$live_birth

results1 = gwqs(abdominal ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$live_birth,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```

## there's a signal only in the positive direction/ negative non-significant
```{r}
chem_mix<- names(dat)[2:33]
offset <- dat$live_birth

results1 = gwqs(cardiac ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$live_birth,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```
## there's a signal only in the positive direction/ negative - significant
```{r}
chem_mix<- names(dat)[2:33]
offset <- dat$live_birth

results1 = gwqs(cranial ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$live_birth,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```

#significant in positive direction
```{r}
chem_mix<- names(dat)[2:33]
#offset <- dat$ped_popln

results1 = gwqs(lymphomas_and_reticuloendothelial_neoplasms ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$ped_popln,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```
#significant in both directions
```{r}
chem_mix<- names(dat)[2:33]
#offset <- dat$ped_popln

results1 = gwqs(malignant_bone_tumors ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$ped_popln,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = TRUE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 > 0) 

summary(results1$fit)
results1$final_weights
```
#significant in negative direction
```{r}
chem_mix<- names(dat)[2:33]
#offset <- dat$ped_popln

results1 = gwqs(neuroblastoma_and_other_peripheral_nervous_cell_tu ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$ped_popln,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```

# other neoplasms did not run in positive/neg due to missing values
#renal sig in positive direction
```{r}
chem_mix<- names(dat)[2:33]
#offset <- dat$ped_popln

results1 = gwqs(renal_tumors ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$ped_popln,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```
#significant in positive
```{r}
chem_mix<- names(dat)[2:33]
#offset <- dat$ped_popln

results1 = gwqs(soft_tissue_and_other_extraosseous_sarcomas ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$ped_popln,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```
# Not significant in either directions
```{r}
chem_mix<- names(dat)[2:33]
#offset <- dat$ped_popln

results1 = gwqs(retinoblastoma ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$ped_popln,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```
# not significant in both directions
```{r}
chem_mix<- names(dat)[2:33]
#offset <- dat$ped_popln

results1 = gwqs(hepatic_tumors ~ wqs, 
               mix_name = chem_mix, 
               offset= dat$ped_popln,
               zero_infl = FALSE, #only for cancer sub-types
               data = dat,
               q=10, validation= 0.6, b = 10, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(results1)
gwqs_scatterplot(results1)
gwqs_summary_tab(results1)
sum(results1$bres$b1 < 0) 

summary(results1$fit)
results1$final_weights
```

######################################################################################
######################################################################################
# Analysis strategy 2
#Estimate WQS adjusted by covariates without constraining; count the number of betas that are positive/negative out of B
# Constraining in the direction that was significant
#positive direction
```{r}
result2 <- gwqs(hepatic_tumors ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln,
               zero_infl = FALSE, #only for sub-types
               data = dat,
               q=10, validation= 0.6, b = 100, 
               b1_pos = TRUE, b1_constr = FALSE, #significant positive direction
               family= "negbin", 
               seed = 2023)

gwqs_barplot(result2)
gwqs_summary_tab(result2)
gwqs_scatterplot(result2)
sum(result2$bres$b1 > 0)  
sum(result2$bres$b1 < 0) 

```

#negative direction
```{r}
result2 <- gwqs(hepatic_tumors ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln,
               zero_infl = FALSE, #only for sub-types
               data = dat,
               q=10, validation= 0.6, b = 100, 
               b1_pos = FALSE, b1_constr = FALSE, 
               family= "negbin", 
               seed = 2023)

gwqs_barplot(result2)
gwqs_summary_tab(result2)
gwqs_scatterplot(result2)
sum(result2$bres$b1 < 0)  
sum(result2$bres$b1 > 0) 

```

#Analysis strategy 3- compare to random subset WQS - Total_can,CNS,Germ cell,malignant bone,renal and soft tissue were positive and significant
#b=500,400,300,250,100 data did not converge
```{r}

result4 = gwqs(cns_and_miscellaneous_intracranial_and_intraspinal ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct,
               mix_name = chem_mix, offset=dat$ped_popln,data = dat, q = 10, 
               validation = 0.6,  b = 500, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123,rs=TRUE, n_var=5) # 5 for set of 18

gwqs_summary_tab(result4) 
gwqs_barplot(result4)
gwqs_scatterplot(result4)
gwqs_weights_tab(result4)
sum(result4$bres$b1 > 0)
```
#Analysis strategy 4- with interaction term- # WQS with interaction - b12 significant (p=0.00000310)
```{r}

result2int = gwqs(total_can ~ wqs*m_pop +vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct,
               mix_name = chem_mix, offset=dat$ped_popln,data = dat, q = 10, 
               validation = 0.6,  b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123)

gwqs_summary_tab(result2int) 
gwqs_barplot(result2int)
gwqs_scatterplot(result2int)
gwqs_weights_tab(result2int)
```

#Analysis strategy 4- with interaction term- # WQS with interaction - b12 Non-significant (p=0.378)
```{r}

result2int = gwqs(total_can ~ wqs*f_pop +vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct,
               mix_name = chem_mix, offset=dat$ped_popln,data = dat, q = 10, 
               validation = 0.6,  b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123)

gwqs_summary_tab(result2int) 
gwqs_barplot(result2int)
gwqs_scatterplot(result2int)
gwqs_weights_tab(result2int)
```

#Analysis strategy: stratified variables in the mixtures- this does not work for my data due to poisson distribution

#Analysis strategy 5: #repeated holdout WQS
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(total_can ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "total_cancer_WQSwtsrh5.csv")

```

#Analysis strategy 5: #repeated holdout WQS
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(cns_and_miscellaneous_intracranial_and_intraspinal ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "cns_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(germ_cell_tumors_trophoblastic_tumors_and_neopla ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "germcell_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS #no valid set of coefficients has been found
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(malignant_bone_tumors ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln,zero_infl = TRUE,# Zero-infl true as had no valid coefficients
               data = dat, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "malig_bone_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS #no valid set of coefficients has been found
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(renal_tumors ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat,
               zero_infl = TRUE,# Zero-infl true as had no valid coefficients
               q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "renal_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS #no valid set of coefficients has been found
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(soft_tissue_and_other_extraosseous_sarcomas ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, 
               zero_infl = TRUE,# Zero-infl true as had no valid coefficients
               q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "soft_tissue_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS 
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(leukemia ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, 
               zero_infl = FALSE,
               q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "leukemia_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS #no valid set of coefficients has been found
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(lymphomas_and_reticuloendothelial_neoplasms ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, 
               zero_infl = TRUE,
               q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "lymphoma_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS #no valid set of coefficients has been found
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(neuroblastoma_and_other_peripheral_nervous_cell_tu ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, 
               zero_infl = TRUE,
               q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "neuroblastoma_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS #no valid set of coefficients has been found
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(retinoblastoma ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, 
               zero_infl = TRUE,
               q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "retinoblastoma_WQSwtsrh5.csv")

```


#Analysis strategy 5: #repeated holdout WQS #no valid set of coefficients has been found
```{r warning=FALSE}
############################### takes FOREVER ##########################
#repeated holdout WQS
# positive direction
result5rh = gwqsrh(hepatic_tumors ~ wqs+vehicle_pct+sing_pent_pct+ssi_snap_pct+
                 m_blk_pct+f_blk_pct+m_hisp_pct+f_hisp_pct+m_heal_ins_pct+
                 f_heal_ins_pct, 
               mix_name = chem_mix, 
               offset=dat$ped_popln, data = dat, 
               zero_infl = TRUE,
               q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = TRUE, family = "negbin", 
               seed = 123, rh=5)
gwqs_summary_tab(result5rh)
gwqsrh_boxplot(result5rh)
gwqs_weights_tab(result5rh)
summary(result5rh$fit)
result5rh$final_weights
result5rh$final_weights$mean_weight = result5rh$final_weights$Estimate

#
write.csv(result5rh$wmat, "hepatic_WQSwtsrh5.csv")

```