---
title: "10_bivariate_maps"
author: "Taiba"
date: "2024-03-24"
output: pdf_document
---

```{r}
library(pacman)
pacman::p_load(tidyverse, janitor, tidycensus, leafsync,ggspatial, mapview, rmarkdown, knitr, stringr, biscale, cowplot, sf, tigris, spdep, tmap)
```

```{r}
ne_counties <- counties(state = "NE", cb = FALSE, resolution = "500k", year = "2019", class="sf") |>
  clean_names()
```


```{r}
index <- read_csv("ne_mapping_dat_final.csv") |> 
  clean_names() |> 
  select(county_fips_code, pca_ln_x_1) |>
  mutate(geoid10 = as.character(county_fips_code),
         geoid10 = str_pad(geoid10, width = 3, side = "left", pad = "0"),
         pca_ln_x_1= pca_ln_x_1*(-1))

out <- read_csv("final_dissrt_analy_data.csv") |>
  clean_names() |>
  select(-c(2:33, 35:43, 49:54, 57:72)) |>
  mutate(geoid10 = str_remove(geoid10, "^31"), 
         geoid10 = str_pad(geoid10, width = 3, side = "left", pad = "0"))
```

#join cancer rates with nebraska shapefile
```{r}
#calculated Annual rate of pediatric cancer incidence per 100000 children
dat_viz<- full_join(index, out, by = "geoid10") |>
  mutate(across(c(cns_and_miscellaneous_intracranial_and_intraspinal:total_can),
                ~ ((./22) / ped_popln) * 100000)) |>
  rename(countyfp = geoid10)

dat_viz_shp<- left_join(ne_counties, dat_viz, by= "countyfp")
```


#============================================================================#
#CNS - bivariate summary
```{r}
data_cns <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = cns_and_miscellaneous_intracranial_and_intraspinal, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_cns |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,27)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#CNS- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_cns, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "CNS cancer rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_cns_pest_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```

#============================================================================#
#============================================================================#

#============================================================================#
#leukemia - bivariate summary
```{r}
data_leukemia <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = leukemia, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_leukemia |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,27)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#leukemia- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_leukemia, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Leukemia rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_leukemia_pest_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```

#============================================================================#
#============================================================================#

#============================================================================#
#lymphoma - bivariate summary
```{r}
data_lymphoma <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = lymphomas_and_reticuloendothelial_neoplasms, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_lymphoma |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,27)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#Lymphoma- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_lymphoma, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Lymphoma rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_lymphoma_pest_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```


#============================================================================#
#============================================================================#
#============================================================================#
#overall cancer - bivariate summary
```{r}
data_overall <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = total_can, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_overall |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,27)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#overall- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_overall, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Overall cancer rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_overall_cancer_pest_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```

############################################################################################
#============================================================================#
#Malignant bone - bivariate summary
```{r}
data_malig_bone <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = malignant_bone_tumors, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_malig_bone |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,27)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#Malignant bone- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_malig_bone, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Malignant bone tumors",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_malig_bone_pest_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```

#######################################################################################################

# Birth defects bivariate maps
```{r}
index <- read_csv("ne_mapping_dat_final.csv") |> 
  clean_names() |> 
  select(county_fips_code, pca_ln_x_1) |>
  mutate(geoid10 = as.character(county_fips_code),
         geoid10 = str_pad(geoid10, width = 3, side = "left", pad = "0"),
         pca_ln_x_1= pca_ln_x_1*(-1))

out <- read_csv("final_dissrt_analy_data.csv") |>
  clean_names() |>
  select(-c(2:33, 35:43, 44:56)) |>
  mutate(geoid10 = str_remove(geoid10, "^31"), 
         geoid10 = str_pad(geoid10, width = 3, side = "left", pad = "0"))
```

#join birth defects rates with nebraska shapefile
```{r}
#calculated Annual rate of birth defects per 100000 children
dat_viz<- full_join(index, out, by = "geoid10") |>
  mutate(across(c(abdominal:total_bd),
                ~ ((./22) / live_birth) * 100000)) |>
  rename(countyfp = geoid10)

dat_viz_shp<- left_join(ne_counties, dat_viz, by= "countyfp")
```


#============================================================================#
#Total birthdefects - bivariate summary
```{r}
data_totaldef <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = total_bd, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_totaldef |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,38)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#total birth defects- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_totaldef, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Overall Birthdefects rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```



```{r}
ggsave("ne_total_bd_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```

#============================================================================#
#============================================================================#

#============================================================================#
#Cardiac defects - bivariate summary
```{r}
data_cardiac <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = cardiac, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_cardiac |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,38)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#Cardiac defects- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_cardiac, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Cardiac defects rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_cardiac_bd_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```

#============================================================================#
#Hepatic defects - bivariate summary
```{r}
data_hepatic <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = hepatic, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_hepatic |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,38)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#hepatic defects- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_hepatic, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Hepatic defects rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_hepatic_bd_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```


#============================================================================#
#renal defects - bivariate summary
```{r}
data_renal <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = renal_genuri, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_renal |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,38)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```

#renal defects- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_renal, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Renal defects rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_renal_bd_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```


#============================================================================#
#cranial birthdefects - bivariate summary
```{r}
data_cranial <- bi_class(dat_viz_shp, x = pca_ln_x_1, 
                      y = cranial, 
                      style = "quantile", 
                      dim = 3)

summary_bi_variate <- data_cranial |>
  filter(bi_class == "3-3" | bi_class == "3-2") |>
    as_tibble() |>
  select(c(5,38)) |>
  group_by(bi_class) %>%
  summarise(names = paste(name, collapse = ", "))
```


#Cranial birth defects- Bivariate Mapping with Biscale
```{r}
# create map
map <- ggplot() +
  geom_sf(data = data_cranial, mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink2", dim = 3) +
  labs(
    title = "",
    subtitle = ""
  ) +
  bi_theme()
```

# creating legends
```{r}
legend <- bi_legend(pal = "GrPink2",
                    dim = 3,
                    xlab = "Pesticide index ",
                    ylab = "Cranial defects rate",
                    size = 0.2)+
  theme(text = element_text(size = 10))
```

# creating final output 
```{r}
# combine map with legend
ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.05, -0.09, 0.2, 0.7)
  #draw_plot(legend, 0.7, .2, 0.2, 0.09)
```

```{r}
ggsave("ne_cranial_bd_bivariate.tiff",
       width = 12, height = 10, dpi = 300)
```
